### 2015-12-17 special-issue-ci/Response to reviewers-r2.tex
including reviewers' questions

\documentclass[preprint]{elsarticle}
\biboptions{round, numbers}
\usepackage[latin1]{inputenc}
%\usepackage[T1]{fontenc}
%\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{color}
%\usepackage{setspace}
\usepackage[hyphens]{url}
\usepackage[english]{babel}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources 
for Low Cost Distributed Evolutionary Computation: Response to Reviewers\' comments}

\noindent
Dear Sirs,\\

We really appreciate the opportunity you have given us for improving
our work. Following the reviewers' requests and suggestions we have
deeply changed the contents, and thus, the quality of the paper.

In the next sections you can find these comments and our responses
to them, along with the changes done.\\

\noindent
Yours sincerely,\\
The authors.

\section{Response to Editors}

\section{Response to Reviewer \#1}

\begin{quote}\textbf{

}\end{quote}

\section{Response to Reviewer \#2}

\begin{quote}\textbf{
}\end{quote}

\section{Response to Reviewer \#3}

\begin{quote}\textbf{
Please, correct the information about i7-4770 processor. Additionally, I would recommend to precise the concept of safety to "user's safety" (or "voluteer's safety") in the characteristic of the framework (page 5), as you still do not address the safety of the system as a whole, at least not at this point.

I would also recommend to include at least a short information that the system utilizes HTML5 Web Workers, as from my point of view it is one of the most important aspects of the framework and to reorganize the subsections in section IV and V.

I would also change the title of the subsection V.A as it begins with ''F15:..'' which is a little bit confusing.

}\end{quote}

\section{Response to Reviewer \#4}

\begin{quote}\textbf{
}\end{quote}

\bibliographystyle{IEEEtran}
\bibliography{geneura}

\end{document}
%%% Local Variables:
%%% ispell-local-dictionary: "english"
%%% End: